---
title: "Stef's Updates"
subtitle: "March 9 - March 14"
date: "`r format(Sys.Date(), '%d %b %Y')`"
number-sections: false
number-offset: 0
toc: true
toc-depth: 3
fig-width: 6
fig-height: 4
df-print: kable
embed-resources: true
format: 
  html: 
    highlight-style: dracula
    code-tools: true
    code-fold: true
    code-summary: "Show the code"
    code-line-numbers: true
    code-link: true # only works if code-line-numbers: false
    toc-float: true
  pptx:
    fig-format: jpeg
    toc-title: "Table of contents"
    knitr: 
      opts_chunk: 
        dev: "png"
    reference-doc: template.pptx
    execute: 
      echo: false
  docx: 
    toc-title: "Table of contents"
    fig-dpi: 600
    execute:
      echo: false     
knitr: 
  opts_chunk: 
    fig.path: "plots/"
    fig.keep: "high"
    dev: ['png', 'pdf']
    dev.args: 
      pdf: 
        encoding: 'WinAnsi'
        useDingbats: false
bibliography: references.bib # if using references, need a .bib file
#csl: gpl.csl # find your preferred citation style at https://www.zotero.org/styles
#css: style.css # optional html styling
crossref:
  fig-prefix: Fig.
  tbl-prefix: Table
  ref-hyperlink: true
editor: source
---

# Code Accessibility

## [Working on GitHub to keep everything in one place]{.underline}

-   Using `r R.version.string`.

    -   Quarto is being used to create slides. This way, the R code for all plots in this presentation can be accessed at all times. The code is embedded in the pptx file, but they are not shown (see below).

-   View the code and download the data used for this presentation in the GitHub repository: <https://github.com/StefTasayco/Lake-Trout-AK.git>

    -   The file name is "Age-weight.qmd"

```{r}
#| label: setup
#| message: false

# load packages
library(ggplot2)
library(dplyr)
library(writexl)
library(janitor)
library(readxl)
library(forcats)
library(RColorBrewer)

# load color-blind friendly palettes
my_paired <- RColorBrewer::brewer.pal(12, "Paired")
my_paired13 <- colorRampPalette(brewer.pal(12, "Paired"))(13)
```

# Lake Trout Age-Weight

**Boxplots of weight at age** capture the range of possible model input values.

This analysis contributes to the approach that we discussed during the last Lake Trout team meeting where:

-   We pick an age that has robust data (10, 20, 30).

    -   Lake trout follow the [von Bertalanffy growth model](https://en.wikipedia.org/wiki/Von_Bertalanffy_function) (VBGM). aka their growth should plateau as they grow older.

    -   Therefore, we expect that age 10 lake trout will have a higher growth and consumption rate than age 30 lake trout.

-   Create design profile (one per age) using median and quartile weights. Run names for this design profile are shown in @tbl-runnames.

```{r}
#| label: tbl-runnames
#| tbl-cap: Run names in each design profile/cohort and their interpretations.
#| tbl-cap-location: bottom

run_names <- read_excel("data/metadata.xlsx", sheet = "age_weight_runnames")
run_names
```

# Data

```{r}
# load data
datafull <- read_excel("data/lake_trout_datafull.xlsx")
```

## 

# References

::: {#refs}
:::